---
title: "Fars Crices Report Overview"
author: "Garvish Raval"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fars Crices Report Overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, echo = FALSE, include = FALSE}
library(farreport)
library(maps)
```

This vignette gives a brief overview of the FARS package functionality. The package contains a set of functions suited for analysis of data from FARS - Fatality Analysis Reporting System. FARS is a nationwide census providing NHTSA, Congress and the American public yearly data regarding fatal injuries suffered in motor vehicle traffic crashes. [[1](http://www.nhtsa.gov/Data/Fatality-Analysis-Reporting-System-(FARS))]

## Package functions

There are three exported functions available to users:

- `fars_read()`
- `make_filename()`
- `fars_read_years()`
- `fars_summarize_years()`
- `fars_map_state()`

Further we give a short description with examples how to use the functions. For the purposes of these examples we will use data from FARS, years 2013-2015, that can be found in the package directory under `\extdata` folder.

## Read FARS data files

The function `fars_read()` provides a way how to read data from a file path provided as an argument. It returns a `tibble` with loaded data:

```{r fars_read_example}
filename <- system.file("extdata/accident_2014.csv.bz2", package = "farreport")
fars_read(filename)
```

## makes FARS data file name from Year

The function `make_filename()` makes datafile name from year provided with pattern "accident_<year>.csv.bz2".

```{r make_filename_example}
make_filename(2014)
```

## Read FARS data file for provided year(s)

The function `fars_read_years()` reads datafile for provided year(s) and reurn data frame in Array format.

```{r fars_read_years_example}
setwd(system.file("extdata", package = "farreport"))
fars_read_years(2014)
fars_read_years(c(2014,2015))
```

## Summarize number of accidents

The next function `fars_summarize_years()` takes a vector or list of years (numeric/integer values) as an argument and if the corresponding files with data are available, it returns a pivot table with number of accidents per year and month. The data files need to be located in the working directory.

```{r fars_summarize_years_example}
setwd(system.file("extdata", package = "farreport"))
fars_summarize_years(2013)
```

## Plot accident locations

The last function available in the package, `fars_map_state()` takes a state index and year as arguments and returns a plot of the state with accident locations. The state indices need to correspond available indeces from the data set. Here are some examples:

```{r fars_map_state_example}
setwd(system.file("extdata", package = "farreport"))
fars_map_state(45, 2015)
```

Note that again, the data file needs to be present in the working directory. Examples are setting working directory to package's extdata which contains sample data files.
